\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}

% From uithesis.sty
\usepackage{listings}
\usepackage{xcolor}
\usepackage{float}
\usepackage{array}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{chains}
\usepackage{pgfplots}

\newcommand{\func}[2]{\text{#1}\left(#2\right)}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Implementation and Evaluation of Multi-Dimensional Defragmentation and Network-Aware Filtering Policies for Kubernetes Descheduler}

\author{
    \IEEEauthorblockN{Naufal Ichsan\IEEEauthorrefmark{1}, Muhammad Raihan Akbar\IEEEauthorrefmark{2}, Matthew Hotmaraja Johan Turnip\IEEEauthorrefmark{3}, and Made Harta Dwijaksara\IEEEauthorrefmark{4}}\\
    \IEEEauthorblockA{Faculty of Computer Science, Universitas Indonesia, Depok 16424, Indonesia}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{1}naufal.ichsan@ui.ac.id}
}

\markboth{Journal Paper Draft}%
{Ichsan \MakeLowercase{\textit{et al.}}: Implementation and Evaluation of Multi-Dimensional Defragmentation}

\IEEEtitleabstractindextext{%
\begin{abstract}
Dynamic Kubernetes clusters accumulate multi-dimensional resource fragmentation that strands capacity, while compute-focused consolidation may disrupt busy workloads or move dependent microservices across degraded network links. This study presents a Kubernetes Descheduler architecture that combines resource defragmentation with actual-usage and network-aware eviction safeguards. The proposed \textit{ResourceDefragmentation} policy evaluates CPU-memory shape using the Resource Imbalance Index (RII), Free-Space Index (FSI), and bin scores, then selects evictions based on feasible predicted scheduler targets. Actual-usage and network-aware filters respectively protect busy Pods and block movements toward latency-degrading destinations. Across five request-based resource scenarios, the policy reclaimed four Nodes in the underutilized scenario and three Nodes in the fragmented scenario. In the heterogeneous-drain scenario, it reduced Mean Absolute RII from the low-20\% range to around 2\% using two evictions. In the controlled single-replica workload, the actual-usage filter protected the busy API Pod by redirecting eviction to an idle candidate, preventing the failures observed when the API Pod was evicted. In the network-aware evaluation, the baseline ResourceDefragmentation behavior balanced compute resources but evicted sensitive workloads, causing latency spikes of up to $\sim$167ms. In contrast, the latency-aware filter intercepted these network-degrading actions, maintained baseline latency at $\sim$8ms, and still reduced Mean Absolute RII from the 50\% range to around 17\% by steering eviction toward non-critical Pods. Overall, the proposed architecture improves resource utilization while protecting active and network-sensitive workloads.
\end{abstract}

\begin{IEEEkeywords}
Kubernetes, Descheduler, Resource Defragmentation, Runtime Resource Usage, Network Latency, Microservices
\end{IEEEkeywords}}

\maketitle
\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\input{src/paper-body/bab1}
\input{src/paper-body/bab2}
\input{src/paper-body/bab3}
\input{src/paper-body/bab4}
\input{src/paper-body/bab5}

\appendices
\section{Resource Defragmentation Balance Filter Algorithm}
\label{appendix:resource_defrag}
\lstinputlisting[
    label={lst:resource_defrag},
]{assets/codes/pseudos/resource-defrag.psu}

\section{Actual Usage Pre-Eviction Filter Algorithm}
\label{appendix:actualusage_algorithm}
\lstinputlisting[
    label={lst:actualusage_algorithm},
]{assets/codes/pseudos/actualusage/unifiedalgo.psu}

\section{Network-Aware Pre-Eviction Filter Algorithm}
\label{appendix:netaware_unified_algorithm}
\lstinputlisting[
    label={lst:netaware_unified_algorithm},
]{assets/codes/pseudos/netaware/unifiedalgo.psu}


\bibliographystyle{IEEEtran}
\bibliography{config/references}

\end{document}
