%
% Halaman Abstract
%
% @author  Andreas Febrian
% @version 2.2.0
% @edit by Ichlasul Affan
%

\chapter*{ABSTRACT}
\singlespacing

\noindent \begin{tabular}{l l p{11.0cm}}
	\ifx\blank\npmDua
		Name&: & \penulisSatu \\
		Study Program&: & \studyProgramSatu \\
	\else
		Writer 1 / Study Program&: & \penulisSatu~/ \studyProgramSatu\\
		Writer 2 / Study Program&: & \penulisDua~/ \studyProgramDua\\
	\fi
	\ifx\blank\npmTiga\else
		Writer 3 / Study Program&: & \penulisTiga~/ \studyProgramTiga\\
	\fi
	Title&: & \judulInggris \\
	Counselor&: & \pembimbingSatu \\
	\ifx\blank\pembimbingDua
	\else
		\ &\ & \pembimbingDua \\
	\fi
	\ifx\blank\pembimbingTiga
	\else
		\ &\ & \pembimbingTiga \\
	\fi
\end{tabular} \\

\vspace*{0.5cm}

Dynamic Kubernetes clusters often accumulate resource fragmentation and multi-dimensional CPU-memory skews, leading to stranded capacity and inefficient infrastructure utilization. While the Kubernetes Descheduler provides a mechanism for post-placement optimization, traditional policies generally optimize compute symmetry through rigid utilization thresholds. However, in modern microservice architectures, standard bin-packing can break network locality, separating dependent services across degraded network links and increasing application latency. In this study, we present a multi-level Kubernetes Descheduler architecture designed to resolve both multi-dimensional resource fragmentation and network performance degradation. To address stranded capacity, we propose a \textit{ResourceDefragmentation} policy that assesses Nodes using average utilization, a CPU-memory bin score, the Resource Imbalance Index (RII), and the Free-Space Index (FSI). The policy predicts feasible scheduler targets before eviction and selects an eviction candidate using the projected CPU-memory bin score of the predicted landing Node, while allowing eviction only when the target remains below a utilization ceiling and is at least as packed as the source. Additionally, we introduce a dual-layer network-aware filtering mechanism that integrates real-time network telemetry to evaluate feasible destinations and block latency-degrading evictions. Across five request-based resource scenarios, the policy reclaimed four Nodes in the underutilized scenario and three Nodes in the fragmented scenario. In the heterogeneous-drain scenario, it reduced Mean Absolute RII from the low-20\% range to around 2\% using two evictions. In the network-aware evaluation, the baseline ResourceDefragmentation behavior balanced compute resources but evicted sensitive workloads, causing latency spikes of up to $\sim$167ms ms. In contrast, the latency-aware filter intercepted these network-degrading actions, maintained baseline latency at $\sim$8ms, and still reduced Mean Absolute RII from the 50\% range to around 17\% by steering eviction toward non-critical Pods. These results show that predicted-target resource defragmentation can improve utilization while explicit network-aware filtering preserves communication integrity.

\vspace*{0.2cm}

\noindent Key words: \\ Kubernetes, Descheduler, Resource Defragmentation, Network Latency, Microservices \\

\setstretch{1.4}
\newpage
