%-----------------------------------------------------------------------------%
\chapter{\babDua}
\label{bab:2}
%-----------------------------------------------------------------------------%
This chapter reviews the theoretical foundation and existing literature relevant to this research. It begins by exploring the underlying architecture of Kubernetes resource management and identifying the core operational challenge: the multi-dimensional resource fragmentation problem. To address this, the concepts of bin-packing and Multi-Criteria Decision Analysis (MCDA), specifically the Technique for Order Preference by Similarity to Ideal Solution (TOPSIS), are introduced. The discussion then focuses on the Kubernetes Descheduler framework, analyzing its current capabilities and inherent limitations. Finally, the chapter examines the need for complementary group eviction policies to support microservice architectures, ultimately setting the stage for the proposed policy-based descheduler framework.


%-----------------------------------------------------------------------------%
\section{Kubernetes Architecture and Resource Management}
\label{sec:k8s_architecture}
%-----------------------------------------------------------------------------%
Kubernetes has emerged as the de-facto standard for container orchestration in cloud-native environments, and the survey by Senjab \textit{et al.} shows how extensive the surrounding scheduling literature has already become~\citep{senjab_survey_2023}. Applications are deployed as Pods, which are then assigned to worker Nodes by the \texttt{kube-scheduler}. Each scheduling attempt is commonly described as a filtering phase followed by a scoring phase: infeasible Nodes are removed first, and the remaining Nodes are ranked according to scheduler plugins and policy configuration~\citep{marchese_tomarchio_2022,senjab_survey_2023}.

This architecture matters because Kubernetes already exposes a limited form of packing awareness at scheduling time. The official \texttt{NodeResourcesFit} plugin supports strategies such as \texttt{MostAllocated} and \texttt{RequestedToCapacityRatio}, both of which explicitly steer new Pods toward denser placements when that is desirable~\citep{k8s_resource_bin_packing}. However, those policies still operate only at admission time for newly scheduled Pods. They do not revisit earlier placements after the workload mix changes.

Kubernetes also supports dynamic workload management through autoscaling. The Horizontal Pod Autoscaler (HPA) and Vertical Pod Autoscaler (VPA) react to utilization signals, while the Cluster Autoscaler adds capacity for unschedulable Pods and removes Nodes only when workloads can be consolidated safely~\citep{cluster_autoscaler_faq}. Because a Pod remains bound to its assigned Node for its lifetime unless it is recreated, the combined effect of workload churn, scaling, and static binding can gradually move the cluster away from an efficient packing state~\citep{larsson_klein_elmroth_2023}.


%-----------------------------------------------------------------------------%
\section{Resource Fragmentation Problem}
\label{sec:resource_fragmentation}
%-----------------------------------------------------------------------------%
The static binding of Pods to Nodes in Kubernetes creates a persistent operational challenge often described as resource fragmentation. Mishra and Bellur formalize this family of problems as Relative Resource Fragmentation (RRF), where a system may still have aggregate free capacity but not in a shape that is useful for future allocations~\citep{mishra_bellur_2016}. In cluster practice, this is often associated with the ``Tetris problem'': workloads of different sizes leave behind mismatched CPU and memory gaps as they scale up and down.

Grandl \textit{et al.} highlight why this problem becomes more severe in multi-resource environments. Modern tasks consume CPU, memory, disk, and network in different proportions, so packing decisions that ignore multi-dimensional fit can create both fragmentation and over-allocation side effects~\citep{grandl_tetris_2014}. Although this thesis concentrates on CPU and memory fragmentation in Kubernetes Nodes, the same argument explains why a one-dimensional notion of utilization is often misleading.

The Kubernetes-specific evidence is particularly important. Larsson \textit{et al.} report that a running cluster may enter a state of resource fragmentation in which hardware cannot be used to its full potential because resources become locked by poor Pod-to-Node mappings~\citep{larsson_klein_elmroth_2023}. Their results support the practical view that fragmentation is not merely an abstract scheduler inefficiency; it materially affects how many Nodes are needed to serve a workload.

Resource fragmentation in this context is therefore best understood as stranded multi-dimensional capacity. A Node can be saturated in one resource dimension while leaving substantial residue in another, and the remaining fragments may be too small or too imbalanced to fit pending Pods. A fragmentation-aware policy must detect these asymmetries and favor actions that restore a more usable capacity profile rather than optimizing only a single utilization percentage.


%-----------------------------------------------------------------------------%
\section{The Bin-packing Problem and Multi-Criteria Decision Analysis}
\label{sec:binpacking_mcda}
%-----------------------------------------------------------------------------%
Addressing resource fragmentation is fundamentally a variation of the classic bin-packing problem: Pods must be arranged onto Nodes so that the number of active Nodes is minimized without violating resource constraints. In a Kubernetes setting, however, this is a multi-resource and time-varying version of bin packing rather than a simple one-dimensional problem.

The Tetris scheduler by Grandl \textit{et al.} is a useful theoretical anchor because it argues that packing based on all relevant resource dimensions reduces fragmentation more effectively than schedulers that rely on only a subset of resource types~\citep{grandl_tetris_2014}. Likewise, Weng \textit{et al.} show in a Kubernetes-based GPU-sharing context that fragmentation can be treated as an explicit optimization target rather than an incidental side effect, and that scheduler decisions can be directed toward the steepest descent of fragmentation~\citep{weng_fgd_2023}. Even though their workload domain is GPU sharing, the paper reinforces the broader principle that fragmentation deserves direct modeling.

To operationalize multi-dimensional decision making in a descheduling pipeline, Multi-Criteria Decision Analysis (MCDA) offers an appropriate abstraction. One prominent MCDA method is Technique for Order Preference by Similarity to Ideal Solution (TOPSIS), introduced by Hwang and Yoon~\citep{topsis_1981}. TOPSIS ranks alternatives according to their distance from an ideal positive solution and an ideal negative solution across multiple criteria. In this thesis, that logic is suitable for ranking eviction candidates because the decision should balance more than one objective, such as reducing node imbalance, improving reclaimability, and avoiding poor locality side effects.

By applying TOPSIS to Kubernetes consolidation, the system can prioritize Pods whose eviction most improves the post-eviction resource profile of a source Node. This is more precise than evicting arbitrary Pods from underutilized Nodes, because the best candidate is the one that most helps transform stranded capacity into a shape that the scheduler can reuse effectively.


%-----------------------------------------------------------------------------%
\section{Kubernetes Descheduler and its Limitations}
\label{sec:k8s_descheduler}
%-----------------------------------------------------------------------------%
Because the \texttt{kube-scheduler} is reactive and only evaluates placement when a Pod is created, it cannot directly repair the long-term fragmentation that accumulates in a running cluster. The Descheduler fills this gap by periodically inspecting the current state, selecting Pods according to policy, and evicting them so their controllers recreate them under the latest cluster conditions~\citep{k8s_descheduler_docs,larsson_klein_elmroth_2023}.

This mechanism makes the Descheduler a natural substrate for consolidation. In particular, policies such as \texttt{HighNodeUtilization} aim to help free lightly loaded Nodes so that the Cluster Autoscaler can remove them, while other policies address topology or affinity violations~\citep{k8s_descheduler_docs}. Larsson \textit{et al.} are especially relevant here because they provide direct empirical support for directed Pod eviction as a way to reduce fragmentation in Kubernetes rather than treating descheduling as a purely heuristic maintenance action~\citep{larsson_klein_elmroth_2023}.

Despite that promise, the standard Descheduler strategies are still coarse for the problem addressed in this thesis. Their decisions are generally driven by configurable thresholds and broad policy categories, not by a fine-grained ranking of which specific Pod eviction would best improve a Node's multi-resource shape. As a result, they offer an operational hook for corrective action but not yet a strong decision model for targeted defragmentation.


%-----------------------------------------------------------------------------%
\section{Group Eviction for Microservices}
\label{sec:group_eviction}
%-----------------------------------------------------------------------------%
Beyond resource utilization metrics, proactive descheduling must also account for the structure of distributed applications. Microservice workloads are composed of communicating components, so placement quality is partly determined by whether strongly interacting services remain in favorable network positions.

Marchese and Tomarchio explicitly address this issue by proposing a Kubernetes scheduler extension that scores Nodes using microservice communication requirements, traffic history, and network latency metrics~\citep{marchese_tomarchio_2022}. Their work is valuable for this thesis not because it solves fragmentation, but because it shows that default Kubernetes placement is incomplete for communication-sensitive applications. Pogonip provides a similar lesson from an edge setting, where queue-centric asynchronous applications benefit from placing communicating components with attention to latency between nodes~\citep{pusztai_pogonip_2021}.

These studies motivate the complementary role of communication locality in this thesis. The proposed work does not attempt a full network-aware scheduler replacement, and it does not claim that every related Pod must move atomically as a hard constraint. Instead, it treats dependency and communication cost as an additional eviction consideration: when two candidate evictions offer similar packing benefits, the policy should prefer the one that is less likely to separate strongly interacting services or worsen their network path.


%-----------------------------------------------------------------------------%
\section{Current Approaches and the Research Gap}
\label{sec:research_gap}
%-----------------------------------------------------------------------------%
Taken together, the literature points to a clear gap. First, fragmentation and packing are already recognized problems in both general cluster scheduling and Kubernetes-specific studies~\citep{grandl_tetris_2014,larsson_klein_elmroth_2023,senjab_survey_2023}. Second, the Kubernetes Descheduler provides a practical control point for post-placement correction, but its default policies do not deeply model which eviction best repairs multi-dimensional stranded capacity~\citep{k8s_descheduler_docs}. Third, communication-aware scheduling research shows that placement quality should not be defined solely by resource utilization when workloads contain strong service dependencies~\citep{marchese_tomarchio_2022,pusztai_pogonip_2021}.

Therefore, the research gap addressed by this thesis is not the absence of descheduling, nor the absence of network-aware placement research in general. The gap is the lack of a Kubernetes-oriented consolidation policy that combines targeted fragmentation reduction with a practical locality-aware safeguard. This motivates a policy-based descheduler that uses stronger ranking logic for eviction decisions while preserving the thesis focus on consolidation as the primary objective.
