%-----------------------------------------------------------------------------%
\chapter{\babLima}
\label{bab:5}
%-----------------------------------------------------------------------------%
%-----------------------------------------------------------------------------%
\section{Conclusion}
\label{sec:conclusion}
%-----------------------------------------------------------------------------%
The Resource Defragmentation policy validates the core fragmentation thread of this research discussed in \ref{sec:researchQuestion}. The synthetic cluster walk-through traces an end-to-end lifecycle from a fragmented initial state through TOPSIS-driven eviction selection to default-scheduler placement of both the freed slot and the evicted Pods' replacements, with each numerical step reproducible from the production code. The walk-through further confirms that runtime usage exposes a class of fragmentation invisible to a request-only signal and that EWMA smoothing damps transient spikes that would otherwise drive unnecessary evictions. The simulator-based scale-out, run on a Borg-derived workload over a heterogeneous twenty-node cluster, reduces the number of unscheduled Pods from \(900\) to \(671\) while issuing only \(31\) evictions, yielding a rescheduling efficiency of \(6.94\) compared with at most \(0.43\) for the official utilization-based and greedy baselines. These results indicate that targeted, multi-criteria eviction selection is more effective than threshold-only or churn-heavy strategies under multi-resource pressure.

The empirical evaluation of the developed Network-Aware Descheduler across three distinct architectural profiles---load-balancing (\textit{LowNodeUtilization}), workload-packing (\textit{HighNodeUtilization} and (\textit{ResourceDefragmentation}))---successfully validates the network-aware thread of this research discussed on \ref{sec:researchQuestion}. Traditional upstream Kubernetes descheduling strategies operate blindly with respect to physical and logical network topologies, consistently sacrificing application performance by shuffling network-critical workloads across high-latency wide-area network links to satisfy raw compute request symmetry. The experimental results demonstrate that the proposed dual-layer network-aware architecture mitigates this structural vulnerability by maintaining a strict balance between resource allocation variance and network locality constraints. In load-balancing scenarios the system restricted compute request variance to a CPU and memory standard deviation of \(14.3\%\), pinned the average cross-service communication latency to \(1.216\,\text{ms}\), and achieved exactly zero pods spilled cross-region, compared with the upstream baseline's \(82.590\,\text{ms}\) network latency penalty. Under severe multi-dimensional skews, the Layer 1 pre-eviction filter intercepted volatile balancing loops, kept network latency capped at \(6.138\,\text{ms}\) versus the baseline's \(98.370\,\text{ms}\), and again held the cross-region pod spill at zero.

A critical systemic limitation was uncovered during \textit{HighNodeUtilization} target filtering testing. When the descheduler attempts to enforce network-aware placement at Layer 2 in isolation, its intent can be entirely subverted by a network-blind default Kubernetes scheduler. Due to standard, density-driven \textit{MostAllocated} scoring rules, pods evicted with the expectation of landing on optimal hosts were frequently misrouted onto degraded nodes because the optimal nodes were already compute-saturated. This confirms that high-precision network locality preservation cannot be solved by the descheduling engine alone; it requires explicit, transactional synchronization between the descheduler and the primary cluster scheduler, an observation that also applies to the Resource Defragmentation feasibility guard, which is request-aware and therefore relies on the scheduler's scoring strategy to keep evicted Pods away from their origin Nodes.

%-----------------------------------------------------------------------------%
\section{Future Works}
\label{sec:future_works}
%-----------------------------------------------------------------------------%
Based on the architectural discoveries and constraints documented throughout this implementation, the following areas are proposed for future research:

\begin{itemize}
    \item \textbf{Coordinated Dual-Engine Scheduling Framework:} Future research should focus on designing a unified extensible network-aware framework for the primary Kubernetes scheduler that can work together with the default descheduler.
    \item \textbf{Adaptive Dynamic-Threshold Filtering Adjustments:} The \textit{NetworkCostEvictor} currently utilizes a static constraint threshold (\textit{minBetterCandidatesPercent}). Developing a feedback loop algorithm that dynamically tightens or loosens enforcement rules based on active cluster-wide resource scarcity or transient network jitter would maximize balancing freedom without risking SLA violations.
    \item \textbf{Production Traffic Volume Weighting:} Integrating service mesh traffic volume data (e.g., from Istio or Linkerd) into the network cost calculations. Weighting point-to-point network latency coefficients by actual cross-service packet throughput (e.g., prioritizing Payment $\leftrightarrow$ Database over less active service pathways) would yield highly efficient, traffic-driven workload alignment.
    \item \textbf{Broader Scenario Coverage for Resource Defragmentation:} The synthetic walk-through and the simulator-based scale-out cover a representative slice of fragmentation patterns (request-only, hidden actual-usage imbalance, transient spikes, and a Borg-derived heterogeneous cluster), but additional scenarios may exercise the policy differently. Future work should sweep more cluster shapes, mixes of CPU-heavy and memory-heavy Pod templates, and non-stationary workloads to surface cases the current evaluation may have missed and to test the Resource Defragmentation policy on configurations not represented here.
\end{itemize}